Rigidity of Expectation: Additional Evidence from Density Forecasts of Professionals and Households

### 2019-06-28 paper/Draft.tex
Finished drafting IR section. Conclusion. Abstracts. Working on stylyzed facts. Tests tables left.

@@

The theories on how agents form expectations in ways deviating from rational expectation(RE) benchmark have proliferated over the past decade. On one hand, these theories built upon different micro foundations produce some similar macro patterns. For instance, the notion of information rigidity, or expectation rigidity could be micro founded by many different theories, but mostly generate sluggish response to new information.  On the other hand, there are important and subtle differences in testable predictions from these theories for both individual forecasts and aggregate moments of the forecasts. This paper extends the current literature in this front by deriving the testable predictions from various theories on uncertainty and empirically test them utilizing the previously ignored probablistic forecast from both professional forecasters survey and household survey. Overall, the paper finds additional evidence from uncertainty consistent with the rigidity models.

The potential contribution lies in the extra insights I may get from using probablistic questions.\footnote{For an insightful survey on the importance of measuring subjective expectations using probablistic surveys, see \citet{manski2004measuring}.} One obvious advantage of density expectations compared to point one is availability of higher moments, i.e. variance of the forecast indicates the dispersion of the forecasts. Along time dimension, the dynamics of conditional relative likelihood is a useful indicator of information gain or forecasting efficiency. Besides, as in a typical signal-extraction model, the prediction that new information reduce conditional variance can be tested only using probability distributions. Although the detailed implementation of these tests involves technical difficulties, the ideas are quite clear.  Lastly, with second moments in individual level expectations, we are allowed to potentialy explore the heterogeneity across agents within one theory. For instance, sticky information typically assume a posson update rate for all agents in the economy. This is partly due to the fact that available data only allows us to recover single parameter of regidity instead of inter-group heterogeneity.

@@

\begin{eqnarray}
\begin{aligned}
\bar \sigma^2_{t+h|t+1} - \bar \sigma^2_{t+h|t} = (1-\lambda)(
\bar \sigma^2_{t+h|t} - \bar \sigma^2_{t+h|t-1}) -\lambda \rho^{2h} \sigma^2 
\end{aligned}
\end{eqnarray}

@@

\subsection{Stylized Facts}

\subsubsection{Different Moments}

Figure \ref{UnceratitnyOtherMoments} plots the population uncertainty asgainst forecast errors and disagreements.

The left  three figures look into the relationship between forecast error and uncertainty. One may expect that  greater ex ante uncertainty implies greater ex post forecast errors.

Another widely made statement is that uncertainty is positively correlated with the level of the inflation. I look into this hypothesis by ploting uncertainty against the realized inflation.

Many literature use disagreements and uncertainty interchangably but they are different objects. From the graph, we can also see in general uncertainty captures different information from disagreements. This is more so in professional forecaster data. There is a more obvious correlation between disgreements and uncertainty.

Once implication from FIRE is that all individuals not only agree on mean but also share same degree of uncertainty. While in fact, we observe long-lasting heterogeneity in uncertainty both from professional forecasters and households. Figure \ref{IQR_Unceratitny} show that there is  not only dispersion of individual forecasts i.e. disagreements, but alslo notable heterogeneity in uncertainty. One stylized fact from these observations is that  the dispersion in ucnertainty seems persistent over time.

@@

\begin{figure}[ht]
	\centering
	\smallskip
	\includegraphics[width=4.5cm]{figures/Inf1yf_CPIAU_varSPFCPIQ.png}
	\includegraphics[width=4.5cm]{figures/SPFCPI_FE_varSPFCPIQ.png}
	\includegraphics[width=4.5cm]{figures/CPI_disg_varSPFCPIQ.png}\\
	\smallskip
	\includegraphics[width=4.5cm]{figures/Inf1yf_PCE_varSPFPCEQ.png}
	\includegraphics[width=4.5cm]{figures/SPFPCE_FE_varSPFPCEQ.png}
	\includegraphics[width=4.5cm]{figures/PCE_disg_varSPFPCEQ.png}\\
	\smallskip	
	\includegraphics[width=4.5cm]{figures/Inf1yf_CPIAU_varSCEM.png}
	\includegraphics[width=4.5cm]{figures/SCE_FE_varSCEM.png}
	\includegraphics[width=4.5cm]{figures/Q9_disg_varSCEM.png}\\
		\caption{Unceratinty and Other Moments}
		\label{UnceratitnyOtherMoments}
\end{figure}

@@

\subsection{Test of Null Hypothesis of Rational Expectation}

@@

\subsection{Additional Evidence from Uncertainty}

@@

\subsection{Evidence from Households}

@@

>>>>

\begin{abstract}
	Surveys on density forecasts of macroeconomic variable that are not available until recent dacade provides one additional moment restriction, uncertainty, for testing and exploring the implications of various theories on expectation formation. This paper first documents the persistent dispersion not only in mean forecast but also in uncertainty about future inflation across economists and households. Second,  utilizing the panel data of individual uncertainty, I provde additional test results that reject the null hypothesis of full-information rational expectation (FIRE). Finally, using externally identified shocks to inflation, I show the dynamic responses of forecasting moments to shocks are broadly consistent with theoreis of expectation rigidity, but in the same time bears the evidence for shock-dependent nature of the rigidity.  This sets the stage for estimating shock-specific regidity using uncertainty as the next step of analysis.

\end{abstract}

\newpage

@@

Theories on how agents form expectations in ways deviating from rational expectation(RE) have proliferated over the past decade. This has been also engined by an incresingly welcoming recognition of the use of surveys of macroeconomists in testing theories empirically.  On one hand, various theories built upon different micro foundations produce some similar macro patterns. For instance, the notion of information rigidity, or expectation rigidity could be micro founded by many different theories, but mostly generate sluggish response to new information.  On the other hand, there are important and subtle differences in testable predictions from these theories for both individual forecasts and aggregate moments of the forecasts.

This paper extends the current literature in this front by deriving the testable predictions from various theories on uncertainty and empirically test them utilizing the previously ignored probablistic forecast from both professional forecasters survey and household survey. Overall, the paper finds additional evidence from uncertainty consistent with the rigidity models. But in the same time, the degree of rigidity apears to be shock-specific and bears the difference to households and professional forecasters.

The contributions of this paper primarily rely upon the additional insights from the forecast uncertainty, that is only available if suveyees are asked to assign their perceived probabilities to a range of variables of the forecast. .\footnote{For an insightful survey on the importance of measuring subjective expectations using probablistic surveys, see \citet{manski2004measuring}.} With density forecasts. the variance of indivdual forecast can be estimated as an indicator of the uncertainty. Across different vintages of the forecast, the revsion in uncertainty is a direct measure of  information gain or the degree of forecasting efficiency. Besides, as in a typical signal-extraction model, the prediction that new information reduce conditional variance can be tested only using probability distributions.

@@

\begin{eqnarray}\label{VarSEEq}
\begin{aligned}
\bar \sigma^2_{t+h|t+1} - \bar \sigma^2_{t+h|t} = (1-\lambda)(
\bar \sigma^2_{t+h|t} - \bar \sigma^2_{t+h|t-1}) -\lambda \rho^{2h} \sigma^2 
\end{aligned}
\end{eqnarray}

@@

\subsection{Stylized facts}

Figure \ref{UnceratitnyOtherMoments} plots the population uncertainty asgainst realized inflation,  forecast errors and disagreements in the first, second and third rows, respectively.

It is widely documented in literature \cite{xxx, xxx} and resonated by anecdotal narrative that high inflation is typically associated with high inflation uncertainty. However, the observable correlation of realized inflation and the directly estimated average uncertainty from both  forecasters and households are, if any, negative during the period between 2007-2019. In particular, the correlation coefficients are -0.1, -0.05 and -0.24 for SPF's  CPI forecast, SPF's PCE forecast and SCE's CPI expectation. This may suggest that during a period of persistently low inflation, the conventiaonl positive redux of inflation and uncertainty is not a good description of the relation of the two.

The middle row of Figure \ref{UnceratitnyOtherMoments} looks into the relationship between forecast error and uncertainty. Although according to our benchmark frameowork in \ref{theory}, there is no mathematical correlation between the forecast errors and uncertainty as the former depends on the realized shocks and the later depends on the volatility of the shock, it is worth checking if in the data a greater ex ante uncertainty implies bigger ex post forecast errors. The correlation coefficent of the two are 0.13, 0.06 and 040 for SPF CPI forecasts, SPF PCE forecasts and SCE households forecasts, resptively. Compared to professional forecasts, households forecasts exhibit such a positive correlation.

 The last row in of Figure \ref{UnceratitnyOtherMoments} explores the relationship between disagreements and uncertainty. Many empriical literature in macroecoomics use disagreements and uncertainty as if they are similar concepts. Partly this confusion in practice is due to difficulty of finding the appropriate measure of this in the first place. But as we have presented in \ref{theory}, the two are distinct concepts. The empricial pattern of the two,  as shown in the plots, also confirms that the two are different objects. The correlation coefficients of the two turn out to be negative, -0.29 and -0.46 for SPF forecasts. In stark contrast, households seem to bear a strong correlation of 0.62 between the two moments.

Although the sheer magnitudes of the differences between professionals' moments and those of houseolds are so big that a direct comparison of the two seems redundant, their respective within-agent-type correlation serves a ready checking device of some statistical consistency. In particular, the positive correlation across households' ex ante uncertainty, ex post forecat errors, and cross-section disagreements cannot be completely reconciled by the framework we set up in \ref{theory}.

The presence of persistent disagreements in expectations has been used as  important stylized evidence against the null of identical expectation assumption embedded in FIRE, for instance, \citet{mankiw2003disagreement}. A similar fact checking can be done with respect to individual uncertaintiy. FIRE predicts individual share equal degree of uncertainty as we clearly seen in Equation \ref{VarREEq}. In contrast, SE predicts that uncertainty of individuals differ in that agents are not equally updated at a point of the time (Equation \ref{VarSEEq}). NI allows for possibility of homogeneity in uncertainty only under sringent condistion of the same precesions of the signals and the same prior of uncertainty. Therefore, precense of dispersion of uncertainty across agents can be read against FIRE.

Figure \ref{IQR_Unceratitny} plots the median forecsts along its 25/75 percentiles on the left column and its counterpart in uncertainty in the righ coumn.  it is clear that not only there is long-lasting dispersion in individual forecasts i.e. disagreements, but alslo notable heterogeneity in uncertainty across agents. 
----------------------------

@@

\begin{figure}[ht]
	\centering
	\smallskip
	\includegraphics[width=4.5cm]{figures/Inf1yf_CPIAU_varSPFCPIQ.png}
	\includegraphics[width=4.5cm]{figures/Inf1yf_PCE_varSPFPCEQ.png}
		\includegraphics[width=4.5cm]{figures/Inf1yf_CPIAU_varSCEM.png}\\
	  	\smallskip
		\includegraphics[width=4.5cm]{figures/SPFCPI_FE_varSPFCPIQ.png}
	\includegraphics[width=4.5cm]{figures/SPFPCE_FE_varSPFPCEQ.png}
		\includegraphics[width=4.5cm]{figures/SCE_FE_varSCEM.png}\\
	\smallskip	
	\includegraphics[width=4.5cm]{figures/CPI_disg_varSPFCPIQ.png}
	\includegraphics[width=4.5cm]{figures/PCE_disg_varSPFPCEQ.png}
	\includegraphics[width=4.5cm]{figures/Q9_disg_varSCEM.png}\\
		\caption{Unceratinty and Other Moments}
		\label{UnceratitnyOtherMoments}
\end{figure}

@@

\subsection{Test of null hypothesis of rational expectation}

@@

\subsection{Additional evidence from uncertainty}

@@

\subsection{Evidence from households}

@@

In summary, three major conclusions can be drawn from this paper.  First, full information rational expectation implicitly assumes agents share the common information and agree on the model that generates the data. Therefore, agents do not only agree on the mean forecast of a macroeconomic variable but also share the same degree of uncertainty. The empirical dispersion of individual uncertainty in future inflation seems to be inconsistent with this simple prediction.

Second, rigidity of expectation from either inattention or underreaction due to noisiness of the signals imply agents do not revise forecasts as efficienctly as they do in FIRE in the face of new information. This can be seen from both a sluggish pattern of forecast revision or uncertainty reduction across differen vintages of the forecast. A weaker test of this spirit uses change of forecasts with fixed horizon. These tests support rigidity models in addition to other null hypotheses of rational expectation including unbiasedness,  non-serial-correlation of forecast error. etc.

Lastly, the directly estimated response of households and professional forecasters' expectation to shocks to inflation suggest that rigidity differs across the shocks. In particular, professional forecasters are more responsive to monetary policy shocks and households are more responsive to oil price shocks. This motivates next-step anlysis of how much resolution of uncertainty can be attributed to the volatility of various shocks. This shows the use of uncertainty.

So far, the dynamics of moments are separately estimated for each moment. Therefore, as a natural extension of the current analysis, one can  undertake a generalized method of moment (GMM) estimation that simultaneously utilize multi-moment restrictions to recover a better estimate of the rigidity parameters.  This will provide a better parameter that can be used in structural macro models that build in rigidity of expectations.
